Creating More Meaningful Explanations for Audio Deepfake Detection

### 2025-01-22 draft.tex
several changes to multiple sections

	with a novel model architecture to generate explanations for fake or real audio classification
	that are potentially more useful to their end user.

@@

	comparison with imperceptible hand-crafted and deep-learning features.
	\newline

@@

	approach using perceptible features for classification in their works, resulting in lackluster
	performance in 2022, with better results using an expanded and improved feature set in 2023.
	\newline
	Li et al. \cite{li_comparative_2022} also approach perceptible features in their work,
	combining them with imperceptible (referred to as physical in the work) features and using
	various neural networks in their experiments. Overall, the combination of perceptible and
	imperceptible features was able to produce the best performance results of the experiments
	tried.

@@

	log-scaled power spectrograms as a feature.
	\newline
	Another example is presented by Haq et al. \cite{haq_multimodal_2023}, who use the changes in
	emotional state as an input feature and represent "unlikely" changes on a graph so that it can
	be understood by an eventual end user.

@@

					estimated based on the maximum power harmonics. This feature is not identical
					to, but is inspired by the use of summarized pitch fluctuations in previous
					research \cite{khanjani_learning_2023}. This feature is also one-dimensional
					with time as the axis. Given that \(H_{i}\) is the set of harmonic frequencies
					at fundamental frequency cycle \(i\) and \(s(h_{i})\) is the power of a given
					harmonic frequency at cycle \(i\), the pitch can be estimated as follows:

@@

			together and pooled into a final output value (in what I will further refer to as the
			terminus). This modular structure allows not only better performance and less memory
			use, due to the fact that fewer transformations are required in input preparation, but
			also allows for more flexibility in the construction of the model, making it easy to
			add and remove features using pre-defined functions depending on feature type. This
			kind of architecture also provides an advantage to the generation of explanations, as
			only the terminus part of the entire must be explained to interpret the importance of
			each input feature. Further explanation is in the following section. Figure
			\ref{fig:model_plot} represents the layout of the model layers.

@@

			The black-box model described in the previous section was trained on the first 10,000
			samples out of the aforementioned dataset, with evaluation being performed on the
			remaining 21,778. Training batches consisted of 100 samples each with every window
			position for each sample and an upper limit of 1,000,000 lines per batch. For each
			batch, two epochs were performed. Because a single sample can have multiple input
			lines, evaluation cannot be performed on a per-line basis, but must instead be
			summarized. Two methods are possible: median and mean, with the difference in results
			being negligible in my experiments. The default threshold is 0.5, meaning that a
			result under 0.5 is a false result and over 0.5 is a true result. This threshold can
			however be adjusted, useful for computing the EER.

@@

		The explanations using the previously described surrogate model are generated with Local
		Interpretable Model-agnostic Explanations (LIME). This section will describe in detail how
		individual explanations are generated at a local level for individual samples, as well as
		how the usefulness of the explanations can be evaluated in general.

@@

			\newline

@@

>>>>

	with a novel model architecture to generate explanations for fake or real audio
	classification, with the goal of singling out features that are useful not only for an
	accurate classification but also for an understandable explanation.

@@

	comparison with imperceptible hand-crafted and deep-learning features in their experiments,
	which were performed on a combination of synthetic and real audio datasets.
	\par

@@

	approach using perceptible features for classification in their works. They trained and tested
	using the dataset of the ADD2022 Challenge \cite{yi_add_2024} and a simple model with a single
	hidden layer, resulting in lackluster performance in 2022, with better results using an
	expanded and improved feature set in 2023.
	\par
	Li et al. \cite{li_comparative_2022} also approach
	perceptible features in their work, combining them with imperceptible (referred to as physical
	in the work) features and using various neural networks in their experiments. Overall, the
	combination of perceptible and imperceptible features was able to produce the best performance
	results of the experiments, which were performed as well on the dataset of the ASV2022
	Challenge \cite{yi_add_2024}.

@@

	log-scaled power spectrograms as a feature, training and testing using the datasets from the
	ASV2019 challenge \cite{wang_asvspoof_2020}, they are able to determine and graphically
	represent areas of importance on the spectrogram, as well as globally summarize the
	SHAP-values.
	\par
	Another example is presented by Haq et al. \cite{haq_multimodal_2023}, who use
	the changes in emotional state as an input feature and represent "unlikely" changes on a graph
	so that it can be understood by an eventual end user. In this case, they achieved their
	results by combining the output of fake video and fake audio classifiers in order to produce a
	final classification for a video sample with audio. Testing against the presidential deepfake
	dataset \cite{sankaranarayanan_presidential_nodate}, they achieve impressive results in
	comparison to the standing benchmark on the dataset at the time.

@@

					estimated based on the maximum power harmonics. This feature is similar to the
					use of summarized pitch fluctuations in previous research
					\cite{khanjani_learning_2023}. This feature is also one-dimensional with time
					as the axis. Given that \(H_{i}\) is the set of harmonic frequencies at
					fundamental frequency cycle \(i\) with \(h_{i} \in H_{i}\) as a frequency of
					the set and \(s(h_{i})\) is the power of a given harmonic frequency at cycle
					\(i\), the pitch can be estimated as follows:

@@

			together, processed though a final model that pools into a final output value (which I
			will further refer to as the terminus model). The modular structure allows not only
			better performance and less memory use, due to the fact that fewer transformations are
			required in input preparation, but also allows for more flexibility in the
			construction of the model, making it easy to add and remove features using pre-defined
			functions depending on feature type. This kind of architecture also provides an
			advantage to the generation of explanations, as only the terminus part of the entire
			must be explained to interpret the importance of each input feature. Further
			explanation is in the following section. Figure \ref{fig:model_plot} represents the
			layout of the model layers.

@@

			The black-box model described in the previous section was trained on the first //TODO//
			samples in the dataset. Evaluation was then performed on the remaining samples in the
			dataset. Training batches consisted of 100 samples each with every window position for
			each sample and an upper limit of 1,000,000 lines per batch. For each batch, two
			epochs were performed. Because a single sample can have multiple input lines,
			evaluation cannot be performed on a per-line basis, but must instead be summarized.
			Two methods are possible: median and mean, with the difference in results being
			negligible in my experiments. The default threshold is 0.5, meaning that a result
			under 0.5 is a false result and over 0.5 is a true result. This threshold can however
			be adjusted, useful for computing the EER.

@@

		The explanations using the previously described surrogate model are generated with the
		Local Interpretable Model-agnostic Explanation (LIME) method \cite{ribeiro_why_2016}. This
		section will describe in detail how individual explanations are generated at a local level
		for individual samples, as well as how the usefulness of the explanations can be evaluated
		in general.

@@

			\par

@@

			In order to determine if the explanations have the potential to be useful to an end
			user, I pursued the goal of summarizing many explanations generated using the
			previously described method together in a meaningful way, in order to make a
			conclusion about the general usefulness of the individual features and the influence
			and potential use of the perceptible features. However, due to a lack of existing
			standardized metrics for the global evaluation of local metrics, I propose two
			metrics, with which I will make conclusions about the usefulness of this method.
